DATA 624 Homework 6

### 2020-04-05 DATA624/HW7.Rmd
More work on HW7

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(gridExtra)
options(scipen=999)
```

@@

(A) Start `R` and use these commands to load the data:
```{r}
library(AppliedPredictiveModeling)
data(permeability)
```

@@

(B) The fingerprint predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.  Filter out the predictors that have low frequencieis using the `nearZeroVar` function from the `caret` package.  How many are left for modeling?

```{r}
library(caret)
length(nearZeroVar(fingerprints))
```

(C) Split the data into training and test sets, pre-process the data, and tune a PLS model.  How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$?

(D) Predict the response of the test set.  What is the test set estimate of $R^2$?

(E) Try building other models discussed in this chapter.  Do any have better predictive performance?

(F)  Would you recommend any of your models to replace the permeability laboratory experiment?

>>>>

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(kableExtra)
library(glmnet)
library(pls)
knitr::knit_hooks$set(inline = function(x) {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(round(x, 2), big.mark = ",")
  }
})
options(scipen = 999)
# Make this reproducible
set.seed(42)
```

@@

### Part A

Start `R` and use these commands to load the data:

```{r}
library(AppliedPredictiveModeling)
data(permeability)
```

@@

### Part B

the `fingerprints` predictors indicate the presense or absense of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure.  Filter out the predictors that have low frequencieis using the `nearZeroVar` function from the `caret` package.  How many are left for modeling?

**`r length(nearZeroVar(fingerprints))`**

### Part C

Split the data into training and test sets, pre-process the data, and tune a PLS model.  How many latent variables are optimal and what is the corresponding resampled estimate of $R^2$?

```{r}
df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
  mutate(y = permeability)
in_train <- createDataPartition(df$y, times = 1, p = .8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

pls_model <- train(
  y ~ ., data = train_df, method = "pls",
  center = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

# Plot model RMSE vs different values of components
plot(pls_model, main=paste("Training Set RMSE Minimized at", pls_model$bestTune$ncomp, "Components"))
```

```{r}
pls_model$results %>%
  filter(ncomp == pls_model$bestTune$ncomp) %>%
  select(ncomp, RMSE, Rsquared) %>%
  kable() %>%
  kable_styling()
```

**The optimal number of components in the PLS model is `r pls_model$bestTune$ncomp`.  This captures `r round(pls_model$results$Rsquared[pls_model$bestTune$ncomp],2)*100`% of the variation in the permeability.**

### Part D

Predict the response of the test set.  What is the test set estimate of $R^2$?

```{r}
# Make predictions
pls_predictions <- predict(pls_model, test_df)
# Model performance metrics
results <- data.frame(Model = "PLS", RMSE = caret::RMSE(pls_predictions, test_df$y), Rsquared = caret::R2(pls_predictions, test_df$y)) 
results %>%
  kable() %>%
  kable_styling()
```

**The $R^2$ is `r round(caret::R2(pls_predictions, test_df$y), 2)`.**

### Part E

Try building other models discussed in this chapter.  Do any have better predictive performance?

#### PCR Model

```{r}
pcr_model <- train(
  y ~ ., data = train_df, method = "pcr",
  center = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

plot(pcr_model, main = paste("Training Set RMSE Minimized at", pcr_model$bestTune, "Components"))

# Make predictions
pcr_predictions <- predict(pcr_model, test_df)
# Model performance metrics
pcr_results <- data.frame(Model = "PCR", RMSE = caret::RMSE(pcr_predictions, test_df$y), Rsquared = caret::R2(pcr_predictions, test_df$y)) 
pcr_results %>%
  kable() %>%
  kable_styling()
```

#### Ridge Regression

```{r}
x <- model.matrix(y ~ ., data=train_df)
x_test <- model.matrix(y ~ ., data=test_df)
rr_cv <- cv.glmnet(x, train_df$y, alpha = 0)
rr_model <- glmnet(x, train_df$y, alpha = 0, lambda = rr_cv$lambda.min)
rr_predictions <- as.vector(predict(rr_model, x_test))
rr_results <- data.frame(Model = "Ridge Regression", RMSE = caret::RMSE(rr_predictions, test_df$y), Rsquared = caret::R2(rr_predictions, test_df$y)) 
rr_results %>%
  kable() %>%
  kable_styling()
```

#### Lasso Regression

```{r}
lr_cv <- cv.glmnet(x, train_df$y, alpha = 1)
lr_model <- glmnet(x, train_df$y, alpha = 1, lambda = lr_cv$lambda.min)
lr_predictions <- as.vector(predict(lr_model, x_test)) 
lr_results <- data.frame(Model = "Lasso Regression", RMSE = caret::RMSE(lr_predictions, test_df$y), Rsquared = caret::R2(lr_predictions, test_df$y))  
lr_results %>%
  kable() %>%
  kable_styling()
```

#### Elastic Net Regession

```{r}
en_model <- train(
  y ~ ., data = train_df, method = "glmnet",
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
)
# Best tuning parameters
en_model$bestTune
```

```{r}

en_predictions <- en_model %>% predict(x_test)
# Model performance metrics
en_results <- data.frame(Model = "Elastic Net Regression", RMSE = caret::RMSE(en_predictions, test_df$y), Rsquared = caret::R2(en_predictions, test_df$y))
en_results %>%
  kable() %>%
  kable_styling()
```

#### Summary

```{r}
pls_model$results %>%
  filter(ncomp == pls_model$bestTune$ncomp) %>%
  mutate("Model" = "PLS") %>%
  select(Model, RMSE, Rsquared) %>%
  as.data.frame() %>%
  bind_rows(pcr_results) %>%
  bind_rows(rr_results) %>%
  bind_rows(lr_results) %>%
  bind_rows(en_results) %>%
  arrange(desc(Rsquared)) %>%
  kable() %>%
  kable_styling()
```

### Part F

Would you recommend any of your models to replace the permeability laboratory experiment?
